\blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}

### 2017-02-02 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates

%%%         %%%
%%% SECTION %%%
%%%         %%%
\section{Introduction}\label{section:Intro}
\medskip
Numerical investigation and modelling of multiphase flows in porous media have attracted the attention of the scientific community over the past 40 years. Characterization of the flow of fluids through porous media serves as the foundation of reservoir engineering studies. Reservoir engineers and hydro-geologists are dealing with complex geometries, irregular boundaries and heterogeneous geological formations. Out of this mixed and complex formations, viscous flow instabilities can occur both within a single phase or when multiple phases flow through porous media. Forecasting the behaviour of multiphase flows or the future response of the system are major challenges that need to be taken into account. Injections of low-viscosity, high-density fluid $1$ (e.g. water) into a reservoir saturated with lighter but more viscous oil is an example where both viscous and gravitational instabilities may occur for two-phase immiscible and incompressible flow. These phenomena have great importance during the oil and gas recovery operations in both mature and new fields. Starting from oil and gas migration of contaminants through groundwater is the opposite problem governed by the same principles. The most complicated of processes include fingering that wil loccur in multiphase and compositional flow. Underground coal gasification is another potential application but more recently on CO$_{\text{2}}$ migration and trapping mechanisms in carbon capture and storage (CCS) operations~\citep{spycher_2003, chen_2006, self_2012, pruess_1990c, white_1981,jiang_2011}.

@@

\section{Problem set-up}\label{problem_set_up}      
Multiphase flow in porous media is described by mass conversation (or transport) equation and Darcy velovities for each of the phase $\alpha$ ($\alpha$ = g, o, w, for gas, oil, and water phases).
Fingering behaviour is expected to be different from single-phase (piston shape front) due to the effect of different peremeabilities. This will change the mobility contrast between two phases for a given \textit{viscosity} ratio. Because of the  different saturation levels among different layers, eventually this will enhance the flow instabilities. In this paper, the test-cases are also based on the discontinuous Galerkin (DG) method to simulate viscous fingering. Roughly speaking, DG-method combines advantages of classical finite volume and finite element methods.

@@

\subsection{Numerical Formulation}
\medskip
Flows in porous media are expressed by the Darcy's law and the saturation equation. At this stage the control volume finite element method is used (CVFEM). Darcy's law for immiscible multi-phase flow may be written in the form:
\begin{equation}\label{e:darcy_eq}
  \mathbf{q}_{\alpha} = -\frac{\mathcal{K}_{{r}_\alpha}\mathbf{K}}{\mu_{\alpha}}\left(\nabla p_{\alpha} - {\mathbf{s}_{u}}_{\alpha} \right),
\end{equation}
where $\mathbf{q}_{\alpha}$ is the $\alpha$-th phase Darcy velocity, $\mathbf{K}$ is the absolute permeability tensor of the porous medium, $\mathcal{K}_{{r}_\alpha}\left(S_{\alpha}\right)$ is the phase relative permeability, which is a function of the phase saturation $S_{\alpha}\left(\mathbf{r},t\right)$. $\mu_{\alpha}$, $p_{\alpha}$, $\rho_{\alpha}$, and $\mathbf{s}_{{u}_\alpha}$ are the phase dynamic viscosity, pressure, density and source term, which may include gravity and/or capillarity, respectively.

@@

\citet{cumming_2011} demonstrated that CVFEM discretisation could also be used to solve Richards' equations (coupled mass conservation and Darcy equation) in heterogeneous media with relatively small computational overhead~\citep[compared with traditional coupled velocity-pressure based formulations, see also][]{cumming_phd2012}. Fluxes over CVs were calculated based upon material properties, whereas the saturation fields were volume-averaged at the interface of the materials, enforcing mass balance as described by \citet{kirkland_1992}.

The conservation of mass along with Darcy's law are the governing equations while variables such as pressure, velocity,saturation and permeability define the behaviour of the flow. Finite difference methods have been applied to discretise the governing equations \citep{Luo2016,Moortgat2016,Hoteit2008} .The objective of this work is to investigate and accurately describe the behaviour of fluid instabilities in compositional flows using a control volume finite element method (CVFEM). Most commercial reservoir simulators use low-order finite-volume/differences methods based on structured grids. Considering the limitations of finite differences formulations, the CVFE The model formulation will use a conservative computational multi-fluid porous media flow model able to exploit mesh adaptivity methods on a fully-unstructured tetrahedral grid. Also two key characteristics that describe the model above are: (a) the hybrid element pair $P_{n}DGP_{n+1}$ which provides discontinuous and piecewise representation for velocity and pressure fields at FE space and (b) the overlapping control volume finite element method (CVFEM)formulation.

@@

%%%         %%%
%%% SECTION %%%
%%%         %%%
\section{Numerical Formulation}\label{section:NumericalFormulation}
The method presented here builds on that reported by \citet{jackson_2015} is known as OCVFEM formulation. The method developed is applicable to $N_{p}$ fluid phases and to the general family of $P_{n+1}(DG)-P_{n}(DG)$ element pair consistent dual pressure-velocity representation in control volume (CV) and finite element (FE) spaces. These element-pairs have a discontinuous $n^{th}$-order polynomial representation for velocity and a continuous representation of pressure. The mass balance (continuity) equations are solved in CV space and a Petrov-Galerkin finite element (FE) method is used to obtain the high-order fluxes on CV boundaries (fig.\ref{fig:fem_elem})  , which are limited to yield bounded fields.

@@

where $\Omega_{CVi}$ and $\Gamma_{CVi}$ are the volume and boundary of CV i respectively, n is normal vector of the CV and $\theta$ varies smoothly between $0.5$ (corresponding to CN) and $1$ (corresponding to BE) to avoid the introduction of spurious oscillations for large grid Courant numbers. The Eq.\ref{global_mass_balance} can be expressed in amtrix form at time level $n+1$ as

%%%         %%%
%%% SECTION %%%
%%%         %%%

\section{Brief Overview of Viscous Instabilities}\label{section:ViscousInstabilities}
%\textcolor{red}{KC: could you add 1-2 short paragraphs on viscous instabilities ... focus on the phenomena rather than a literature review, that may take place in the Intro Section}
As in many areas of fluid mechanics, the behaviour of the flow for conditions that exceed critical limits, are those that occur under highly critical conditions. Under this assumption, viscous fingering is no exception. It is a viscous driven hydrodynamic instability that results the collapse of the uniform interface between the fluids of different viscosity. This type of instability can be found across many research areas, disciplines and scales. Starting from small scale flow instabilities in Hele-Shaw cells that can provide the basic mathematical model, to coupled chromatographic separation systems for chemical and medical applications and reaching up to large scale reservoir systems.

@@

%%%         %%%
%%% SECTION %%%
%%%         %%%
\section{Model Validation}\label{section:ModelValidation}

%%%         %%%
%%% SECTION %%%
%%%         %%%
\subsection{Bucklet-Leverett Equation}\label{section:BuckletLeverettEquation}
\textcolor{red}{JG: text and simulation of P1DGP2, P1DGP1DG fixed and adaptive mesh -- quantitative analysis}\\
\textcolor{blue}{KC: do i have to mention something here?\\}

Models of water and another fluid, such as air, flowing through a
porous material, are used throughout science and
engineering.  Furthermore, individual chemical and biological
substances being transported in the fluids are often fundamentally
important. The mathematical models for these processes take several forms: scalar nonlinear parabolic equations (Richards' equation), scalar hyperbolic equations (Buckley-Leverett equation), systems of degenerate nonlinear parabolic systems (two-phase and compositional flow), and free boundary problems (Muskat's model). These are macroscopic continuum models that govern the distribution and evolution of quantities that are {\em averages} of the densities, pressures, and velocities arising in classical fluid mechanics. The basis of porous media models in microscopic physics is, however, incomplete: many observed phenomena such as history-dependent behavior and various other scale effects have not been satisfactorily modeled, particularly for multiphase,
multicomponent systems at the scale required for environmental
studies.

Modeling of two-phase immiscible flows in porous media is
important for many industrial applications, such as oil recovery and carbon sequestration. Mathematical conceptualizations for such flows include the Buckley–Leverett model, which, in its simplest form, is represented by a nonlinear advection equation for the relative saturation of fluid phases.This equation models the two phase flows that arise oil-recovery problems and physically
represents a mixture of oil and water through the porous medium.
In conservative form the flux function for this
problem is given by:

\begin{equation}
f(u) = \frac{u^2}{u^2 + \nu(1-u)^2}. \label{buckley}
\end{equation}
Here $\nu$ is viscosity ratio and $u$ represents the saturation of water and
lies between $0$ and $1$.

\subsection{Four Regions}\label{section:4Regions}
\textcolor{blue}{KC \textcolor{red}{WR}: text and simulations -- qualitative analysis}\\

\section{Viscous Instabilities}\label{section:ViscousInstabilities}

@@

%%%         %%%
%%% SECTION %%%
%%%         %%%
\section{Brief Overview of Viscous Instabilities}\label{section:ViscousInstabilities}
%\textcolor{red}{KC: could you add 1-2 short paragraphs on viscous instabilities ... focus on the phenomena rather than a literature review, that may take place in the Intro Section}
As in many areas of fluid mechanics, the behaviour of the flow for conditions that exceed critical limits, are those that occur under highly critical conditions. Under this assumption, viscous fingering is no exception. It is a viscous driven hydrodynamic instability that results the collapse of the uniform interface between the fluids of different viscosity. This type of instability can be found across many research areas, disciplines and scales. Starting from small scale flow instabilities in Hele-Shaw cells that can provide the basic mathematical model, to coupled chromatographic separation systems for chemical and medical applications and reaching up to large scale reservoir systems.

@@

%%%         %%%
%%% SECTION %%%
%%%         %%%
\section{Model Validation}\label{section:ModelValidation}

%%%         %%%
%%% SECTION %%%
%%%         %%%
\subsection{Bucklet-Leverett Equation}\label{section:BuckletLeverettEquation}
\textcolor{red}{JG: text and simulation of P1DGP2, P1DGP1DG fixed and adaptive mesh -- quantitative analysis}\\
\textcolor{blue}{KC: do i have to mention something here?\\}

Models of water and another fluid, such as air, flowing through a
porous material, are used throughout science and
engineering.  Furthermore, individual chemical and biological
substances being transported in the fluids are often fundamentally
important. The mathematical models for these processes take several forms: scalar nonlinear parabolic equations (Richards' equation), scalar hyperbolic equations (Buckley-Leverett equation), systems of degenerate nonlinear parabolic systems (two-phase and compositional flow), and free boundary problems (Muskat's model). These are macroscopic continuum models that govern the distribution and evolution of quantities that are {\em averages} of the densities, pressures, and velocities arising in classical fluid mechanics. The basis of porous media models in microscopic physics is, however, incomplete: many observed phenomena such as history-dependent behavior and various other scale effects have not been satisfactorily modeled, particularly for multiphase,
multicomponent systems at the scale required for environmental
studies.

Modeling of two-phase immiscible flows in porous media is
important for many industrial applications, such as oil recovery and carbon sequestration. Mathematical conceptualizations for such flows include the Buckley–Leverett model, which, in its simplest form, is represented by a nonlinear advection equation for the relative saturation of fluid phases.This equation models the two phase flows that arise oil-recovery problems and physically
represents a mixture of oil and water through the porous medium.
In conservative form the flux function for this
problem is given by:

\begin{equation}
f(u) = \frac{u^2}{u^2 + \nu(1-u)^2}. \label{buckley}
\end{equation}
Here $\nu$ is viscosity ratio and $u$ represents the saturation of water and
lies between $0$ and $1$.

\subsection{Four Regions}\label{section:4Regions}
\textcolor{blue}{KC \textcolor{red}{WR}: text and simulations -- qualitative analysis}\\

@@

%%%         %%%
%%% SECTION %%%
%%%         %%%
\section{Viscous Instabilities}\label{section:ViscousInstabilities}

An accurate finite-element simulator with an adaptive and unstructured 2D grid was used in this study of viscous 
instabilities. An unfavourable - mobility-ratio displacement coupled with the heterogeneity of the porous medium results in viscous instability (fingering),which in turn affects the displacement efficiency of im/miscible processes as well as sweep efficiency and breakthrough time. Although numerical investigations 
in the last few years have remarkably improved the physical 
understanding of viscous fingering, some questions still remain 
unanswered.\\

\begin{equation}
B^T M_{\sigma} ^{-1} C{{\underline{p}}^{n+1}} = {\underline{s}_p ^{n+1}} - B^T M_{\sigma} ^{-1} {{\underline{s}}^{n+1}} 
\label{pressure eq}
\end{equation}

The computationally demanding effort to solve the pressure matrix equation (arising from the discontinuous FEM formulation) is achieved using a multigrid-like approach.

The original geometry of the system is an orthogonal 2D domain where the  heterogeneity is imposed by introducing 4 distinct regions of four regions of different absolute permeabilities (K) as can being seen in fig.\ref{fig:1} and fig. \ref{fig:subim6}.

The Richard equation (non-linear advection equation) for relative saturation fluid phases representing two phases immiscible flows in porous media has been simulated. The fluid $1$ or \textit{injected fluid} is injected a a velocity of $u_1 \phi = 1$. On the outlet, or right hand side boundary, the pressure level is set to zero and all the remaining boundary are naturally applied. All the simulations discussed in this section used the overlapping control volume FEM with a piecewise linear variation of the velocity with each element and quadratic pressure ($P_1 DG-P_2$). Saturation is collocated at the  pressure nodes and although it is calculated using a CV formulation, a FEM interpolation is used to form the higher order fluxes. There is no flow across the lower and upper boundaries (no-slip conditions) and the initial velocity of the fluid 2 or displaced fluid is set to zero, while the gravity is assumed negligible.

%%%         %%%
%%% SECTION %%%
%%%         %%%
\subsection{Impact of Permeability Field}\label{section:PermeabilityAnalysis}

The higher the permeability of a medium, the easier it is for flow to propagate through it. Therefore, fluid will flow faster through the higher permeability region. Hence, in Case 1, it is expected that the injected fluid will flow fastest in the top layer which has the highest permeability. The second layer's permeability is higher than that of the bottom layer, so also, its front should be expected to move ahead of that of the bottom layer. The high permeability zone within the lower layer may result in some interesting flow characteristics at the horizontal flow boundary (the top and bottom boundary) depending on the difference between the two permeabilities. If this difference is sufficiently large enough, the fluid from the low permeability region will tend to cross into the high permeability region and flow forward from there. This will result in a region of reduced flow at these boundaries just beyond the deviation from normal flow and may also cause a stationary point at these boundaries within the flow.

\medskip
\textcolor{red}{WR: results }

%%%         %%%
%%% SECTION %%%
%%%         %%%
\subsection{Investigation of Viscosity (Mobility? I think its better if i keep as Mobility Ratio) Ratio}\label{section:MobilityAnalysis}
\textcolor{red}{KC: would you like to add a paragraph on the expected impact of MR in the flow?}\\
\medskip
\textcolor{red}{WR and KC: results}\\
%------------------------KC------------------------------------------

During the simulations performed the Viscous fingering phenomenon has been observed. Viscous fingering  is highly related with the mobility, which describes how easily a fluid moves into a porous media and can be expressed through the mobility ratio (\textit{MR}). The ratio between the mobility of the displacing phase to the mobility of the displaced.

It generally accepted \textcolor{red}{can we say that...?} ( Saffman and Taylor, $1958$, Chouke et al., $1959$, Peters,$1979$ and Homsy,$1987$) that the higher the mobility ratio the easier is for viscous fingering to appear and more specific, for fingers to start to develop and grow. In the simulations above an $MR=10$ is used. This ration combined with the use of unstructured adaptive mesh makes the fingers more visible, easier to detect them. For cases where the is less than 5, ($MR \le 5$) we expect no difference as there is no significant impact. In the extreme case of $MR=1$ we can clearly see the relation between finger development and the zones of different porosity.      
%------------------------KC------------------------------------------

\pagebreak

%%%         %%%
%%% SECTION %%%
%%%         %%%
\section{Final Remarks}\label{section:conclusions}
%\textcolor{red}{KC and WR: do you want to have a bash on this ?}\\ 
%------------------------KC------------------------------------------

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{section:Intro}
\medskip
Numerical investigation and modelling of multiphase flows in porous media have attracted the attention of the scientific community over the past 40 years. Characterization of the flow of fluids through porous media serves as the foundation of reservoir engineering studies. Reservoir engineers and hydro-geologists are dealing with complex geometries, irregular boundaries and heterogeneous geological formations. Out of this mixed and complex formations, viscous flow instabilities can occur both within a single phase or when multiple phases flow through porous media. Forecasting the behaviour of multiphase flows or the future response of the system are major challenges that need to be taken into account. Injections of low-viscosity, high-density fluid $1$ (e.g. water) into a reservoir saturated with lighter but more viscous oil is an example where both viscous and gravitational instabilities may occur for two-phase immiscible and incompressible flow. These phenomena have great importance during the oil and gas recovery operations in both mature and new fields. Starting from oil and gas migration of contaminants through groundwater is the opposite problem governed by the same principles. The most complicated of processes include fingering that wil loccur in multiphase and compositional flow. Underground coal gasification is another potential application but more recently on CO$_{\text{2}}$ migration and trapping mechanisms in carbon capture and storage (CCS) operations~\citep{spycher_2003, chen_2006, self_2012, pruess_1990c, white_1981,jiang_2011}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Problem set-up}\label{problem_set_up}      
Multiphase flow in porous media is described by mass conversation (or transport) equation and Darcy velovities for each of the phase $\alpha$ ($\alpha$ = g, o, w, for gas, oil, and water phases).
Fingering behaviour is expected to be different from single-phase (piston shape front) due to the effect of different peremeabilities. This will change the mobility contrast between two phases for a given \textit{viscosity} ratio. Because of the  different saturation levels among different layers, eventually this will enhance the flow instabilities. In this paper, the test-cases are also based on the discontinuous Galerkin (DG) method to simulate viscous fingering. Roughly speaking, DG-method combines advantages of classical finite volume and finite element methods.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Numerical Formulation}
\medskip
Flows in porous media are expressed by the Darcy's law and the saturation equation. At this stage the control volume finite element method is used (CVFEM). Darcy's law for immiscible multi-phase flow may be written in the form:
\begin{equation}\label{e:darcy_eq}
  \mathbf{q}_{\alpha} = -\frac{\mathcal{K}_{{r}_\alpha}\mathbf{K}}{\mu_{\alpha}}\left(\nabla p_{\alpha} - {\mathbf{s}_{u}}_{\alpha} \right),
\end{equation}
where $\mathbf{q}_{\alpha}$ is the $\alpha$-th phase Darcy velocity, $\mathbf{K}$ is the absolute permeability tensor of the porous medium, $\mathcal{K}_{{r}_\alpha}\left(S_{\alpha}\right)$ is the phase relative permeability, which is a function of the phase saturation $S_{\alpha}\left(\mathbf{r},t\right)$. $\mu_{\alpha}$, $p_{\alpha}$, $\rho_{\alpha}$, and $\mathbf{s}_{{u}_\alpha}$ are the phase dynamic viscosity, pressure, density and source term, which may include gravity and/or capillarity, respectively.

@@

\citet{cumming_2011} demonstrated that CVFEM discretisation could also be used to solve Richards' equations (coupled mass conservation and Darcy equation) in heterogeneous media with relatively small computational overhead~\citep[compared with traditional coupled velocity-pressure based formulations, see also][]{cumming_phd2012}. Fluxes over CV's were calculated based upon material properties, whereas the saturation fields were volume-averaged at the interface of the materials, enforcing mass balance as described by \citet{kirkland_1992}.The conservation of mass along with Darcy's law are the governing equations while variables such as pressure, velocity,saturation and permeability define the behaviour of the flow. Finite difference methods have been applied to discretise the governing equations \citep{Luo2016,Moortgat2016,Hoteit2008} .

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Formulation}\label{section:NumericalFormulation}
The method presented here builds on that reported by \citet{jackson_2015} is known as OCVFEM formulation. The method developed is applicable to $N_{p}$ fluid phases and to the general family of $P_{n+1}(DG)-P_{n}(DG)$ element pair consistent dual pressure-velocity representation in control volume (CV) and finite element (FE) spaces. These element-pairs have a discontinuous $n^{th}$-order polynomial representation for velocity and a continuous representation of pressure. The mass balance (continuity) equations are solved in CV space and a Petrov-Galerkin finite element (FE) method is used to obtain the high-order fluxes on CV boundaries (fig.\ref{fig:fem_elem})  , which are limited to yield bounded fields.

@@

where $\Omega_{CVi}$ and $\Gamma_{CVi}$ are the volume and boundary of CV i respectively, n is normal vector of the CV and $\theta$ varies smoothly between $0.5$ (corresponding to CN) and $1$ (corresponding to BE) to avoid the introduction of spurious oscillations for large grid Courant numbers. The Eq.\ref{global_mass_balance} can be expressed in amtrix form at time level $n+1$ as
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
\section{Brief Overview of Viscous Instabilities}\label{section:ViscousInstabilities}
%\textcolor{red}{KC: could you add 1-2 short paragraphs on viscous instabilities ... focus on the phenomena rather than a literature review, that may take place in the Intro Section}
As in many areas of fluid mechanics, the behaviour of the flow for conditions that exceed critical limits, are those that occur under highly critical conditions. Under this assumption, viscous fingering is no exception. It is a viscous driven hydrodynamic instability that results the collapse of the uniform interface between the fluids of different viscosity. This type of instability can be found across many research areas, disciplines and scales. Starting from small scale flow instabilities in Hele-Shaw cells that can provide the basic mathematical model, to coupled chromatographic separation systems for chemical and medical applications and reaching up to large scale reservoir systems.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Viscous Instabilities}\label{section:ViscousInstabilities}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Overview of Viscous Instabilities}\label{section:ViscousInstabilities}
\medskip
As in many areas of fluid mechanics, the behaviour of the flow for conditions that exceed critical limits, are those that occur under highly critical conditions. Under this assumption, viscous fingering is no exception. It is a viscous driven hydrodynamic instability that results the collapse of the uniform interface between the fluids of different viscosity. This type of instability can be found across many research areas, disciplines and scales. Starting from small scale flow instabilities in Hele-Shaw cells that can provide the basic mathematical model, to coupled chromatographic separation systems for chemical and medical applications and reaching up to large scale reservoir systems.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@
